COVID-19 Disparities in Quebec (Canada)

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/02-body-of-litterature.Rmd
Multiple changes on Bookdown

---
title: 'COVID-19 Disparities in Quebec (Canada): A Comparative Analysis of Exposure
  Risks, Consequences, and Access to Resources among Immigrants and Natives'
author:
- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
- Département de Sociologie, Université du Québec à Montréal (UQAM)
- Georges NGalé Tchango (Département de démographie, Université de Montréal)
- Solène Lardoux (Département de démographie, Université de Montréal)
- Nima Zahedinameghi (Département de Sociologie, UQAM)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    keep_md: yes
  html_document:
    toc: yes
    toc_depth: 3
  template: svm-latex-ms2.tex
  pdf_document: default
  citation_package: natbib
bibliography:
- covid_Nima.bib
- Covid_Canada.bib
- Covid_Family.bib
---
<!--
# Journal: family issues - SAGE
https://journals-sagepub-com.proxy.bibliotheques.uqam.ca/author-instructions/JFI

https://fishandwhistle.net/post/2020/getting-started-zotero-better-bibtex-rmarkdown/
https://www.zotero.org/styles?fields=sociology
-->

# Abstract

The immigrant population is one of the most vulnerable in Canada, with a relatively high risk of unemployment, a low social support system, and linguistic barriers. However, their experiences with public health interventions during the COVID-19 pandemic have not been adequately explored. The purpose of this article is to analyze Covid-19 exposure risk and its family implications for both natives and immigrants in the province of Quebec. Additionally, we analyze immigrants' access to government aid and compare it to their native-born counterparts. The study consists of survey data collected from 1,000 immigrants and 500 non-immigrants Quebecers between August and November 2020. Our descriptive and exploratory analysis reveals significant differences between the two groups regarding exposure risk, family consequences, and access to resources. We conclude that the current pandemic exacerbates health disparities in the province, with particularly dire repercussions for immigrants’ families. We also discuss several socio-cultural strategies to mitigate inequality and recommend policies to address the needs of vulnerable populations.

**Keywords**: Immigrants, exposure risk, government aid, health disparities, COVID-19, family disruptions

<!--

Using  script in Rmarkdown

http://zevross.com/blog/2014/07/09/making-use-of-external-r-code-in-knitr-and-r-markdown/

Managing my Zotero

https://forums.zotero.org/discussion/35712/removing-pdfs-to-free-up-space

https://www.zotero.org/support/sync

https://www.zotero.org/support/preferences/sync#file_syncing

https://www.zotero.org/support/sync#webdav

https://www.zotero.org/support/preferences/advanced#linked_attachment_base_directory

http://zotfile.com/

# Abstract

A ce jour, la pandémie du coronavirus a touché près de 100000 personnes au Québec et entrainé plus de 6000 décès. Cela fait de la province l'une des plus touchées au Canada et l'une des régions les plus affectées au monde. Bien que la majorité de ces décès ait eu lieu dans les maisons pour personnes âgées, et donc a touché plus cette tranche de la population, son influence globale sur les individus n'est pas à négliger. En effet, outre les conséquences directes de la pandémie, les mesures prises par les différents gouvernements à travers le monde ne sont pas sans conséquences sur les familles. Il est démontré que ses conséquences affectent négativement en priorité les couches les plus défavorisées de la population. Les immigrants se concentrent dans bien des cas dans cette couche de la population. Cette étude vise ainsi à analyser l'influence de la pandémie sur les immigrants et leurs enfants au Québec, à examiner les facteurs socio-culturels qui permettent d'atténuer son impact négatif.

To date, the coronavirus pandemic has affected nearly 100,000 people in Quebec and caused more than 6,000 deaths. This makes the province one of the most affected in Canada and one of the most affected regions in the world. Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked. In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are not without consequences for families. It has been shown that its consequences primarily affect the most disadvantaged segments of the population negatively. Immigrants are in many cases concentrated in this segment of the population. This study thus aims to analyze the influence of the pandemic on immigrants and their children in Quebec, to examine the socio-cultural factors that make it possible to mitigate its negative impact.

À soumettre au Canadian Population Studies

Autre projet possible: intégration des immigrants et covid19 au Québec

**Keywords** :

-->

# Introduction / background

The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and health disparities in high-income countries, including Canada. One of the most vulnerable segments of these affluent societies is the immigrant population, that has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult. The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19]. These immigrants face language barriers, low social capital, housing insecurity, and severe poverty, rendering them particularly susceptible to the effects of the current pandemic. Immigrants are also more likely to be employed as essential workers where social distancing, sanitary precautions, and self-isolation, measures put in place to mitigate the pandemy, are challenging [@reid2021Migrant; @yi2021Health; @zhang2021COVID19]. More importantly, migrants may encounter various forms of discrimination, leading to social exclusion and restricted access to healthcare and social services [@spiritus-beerden2021Mental]. It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec. In doing so, this article will reveal a better understanding of how the current pandemic aggravates existing inequalities and challenges in the province for immigrants and their descendants.

<!-- The province of Quebec has one of the highest concentrations of immigrants in Canada, including permanent residents, temporary residents, and refugees [@governmentofcanada2017Daily; @immigration2005Recent]. In 2016, 1.2 million immigrants were living in Quebec, an increase from 8% in 1990 to 13.7% of the population in 2016 (Institut de la statistique du Québec, 2020). <!--As of 2019, the increase in net new residents (60,700 versus 42,200 in 2018) and lower interprovincial migration losses (-1,500 versus -6,100 in 2018) contributed to Quebec's population growth. -->Since the start of the pandemic, most of the confirmed cases were identified in Montreal and areas with high immigrant populations [@miconi2021Ethnocultural; @shields2020COVID19]. The coronavirus pandemic has infected nearly 450,000 people in Quebec and has resulted in more than 11,000 deaths (https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec, consulted on octobre 29, 2021). This makes it the worst affected province in Canada and one of the worst affected regions in the world during the first wave of the pandemic. Although the majority of these deaths took place in elderly's homes, and therefore affected more this segment of the population, its overall influence on individuals is not to be overlooked. It has been shown that, on the one hand, immigrants and their families have a higher risk of being exposed to the pathogen, and, at the same time, they are more vulnerable to racism and discrimination and face barriers to accessing social support [@2021CRARR] and information, especially  for those who didn't speak or understand the two official langages. <!--In addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families. -->Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries. This differentiated effect can be explained in different ways. On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health. Likewise, immigrants often work in jobs where social distancing is difficult to enforce. On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently. The complete shutdown of tye economy with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies. But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.

Despite the growing focus on the differential impacts of the COVID-19 on immigrants and non-immigrants segment of the population, there is still very little knowledge about how the pandemic and its related public policy impact the lives of immigrants and their families. The existing health datasets (largely used to assess this effect) are not necessarily representative of the immigrant population and do not offer an accurate picture of the complex social status and health outcomes among this group [@choi2021Studying; @gagnon2021Immigration; @khan2015Visible]. In most cases, these datasets lack data on the residence length of immigrants and combine newcomers with long-term settlers [@edge2013Discrimination; @hayward2021Clinical]. As a result, they may not allow a detailed assessment of the nature and extent of risk exposure in sub-segments of populations, as well as whether interventions are reaching those most at risk. In addition, the family-level impacts of the pandemic are still poorly understood. While a body of research has explored the association between the pandemic and its potential effects on intimate partner violence and marital dissolution [@bhatnagar2021Navigating; @koshan2020COVID19; @morrison2021Chapter; @prime2020Risk], few studies have examined negative familial repercussions in the context of immigration status. There is ample evidence that sustained family relationships greatly enhance immigrants' lives on several fronts, including fostering economic prosperity [@marger2001Sociala], psychological well-being [@beiser2002Povertya], social network development [@evansluong2017Familya], and educational success for children [@munroe-blum1989Immigranta]. The family is, therefore, a key component of long-term social mobility and economic success of immigrants.

<!--Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce its impact. These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000. On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably. While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life. How these measures make a difference in the face of the pandemic remains to be seen. -->Based on an original biographic survey collected around the first wave of the pandemic in august-november 2020 on immigrants and the native population of Québec, this study mainly aims to understand the family exposure, consequences, and responses of the pandemic on immigrant and native Quebec populations. We assess a whole range of different measures of each outcome. We divided the consequences in three dimensions, 1) access to ressources, 2) ability to perform duties and 3) employment and income. We seek also to analyse in what ways these outcomes differ according to the immigrants length of residence. By doing so, this study provide a clear and global picture of the situation of different population groups in Quebec in the face of the COVID-19. <!-- To respond to these questions, we will use data collected around the first wave of the pandemic in august-november 2020. Our study aims to fill the gap by analyzing group-specific differences in exposure risk, family consequences, and access to resources among immigrants and Quebecers. We also seek to shed light on several socio-cultural strategies to mitigate inequality and recommend relevant policy solutions. The detailed findings will help policymakers address health inequalities and improve equity in Quebec while developing adequate measures and programs. -->

<!-- Introduction  -->

<!-- Like the rest of Canada, Quebec has experienced a substantial increase in its immigration in recent years. The immigrant population in the province thus increased from 8% in the 1990s to 13.7% of the population, or 1.2 million individuals in 2016 (Institut de la statistique du Québec, 2020). This immigration is also unique in that it has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult. The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19]. It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec. -->

<!-- The coronavirus pandemic has affected nearly 100,000 people in Quebec and has resulted in more than 6,000 deaths. This makes it the worst affected province in Canada and one of the worst affected regions in the world. Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked. In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families. Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries. -->

<!-- This differentiated effect can be explained in different ways. On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health. Likewise, immigrants often work in jobs where social distancing is difficult to enforce. On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently. The complete shutdown of the company with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies. But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse. -->

<!-- Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce the impact of the pandemic. These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000. On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably. While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life. How these measures make a difference in the face of the pandemic remains to be seen. -->

<!-- This study mainly aims to understand the family consequences of the pandemic on immigrant and native Quebec populations. Therefore, we will seek to analyse in what ways its consequences differ according to the region of origin of immigrants, length of residence, age and sex. -->

@@

# Theoretical framework 
Family as a unit of analysis can be particularly fruitful when studying health inequalities in the context of a societal-wide crisis. Through the lens of the COVID-19 pandemic, the following points illustrate why exploring family dynamics provides additional insight into the underlying mechanisms of health disparities. First, the family is the most fundamental unit of social organization. It facilitates access to resources and the transmission of social and psychological capital from one generation to the next. Therefore, a better analysis of how crises and adverse life events affect people requires a better understanding of how families relate to each other and interact with external forces. Second, during public health emergencies, families may be thrown into different structures due to unprecedented financial, mental, and physical challenges. The risk of family dissolution, for example, can be heightened when members must spend unusually long periods together at home and have limited opportunities to engage with friends and the wider community. Conversely, families can develop closer bonds through self-sacrifice, shared responsibilities, and increased care and support. As the dependency status of seniors increases, the elderlies are also more likely to be integrated into formerly nucleus families. We posit that the extent to which a family can adapt its functions to meet new challenges is a predictor of the well-being of its members and the resilience of its broader population. Third, the family is a crucial link to social institutions, such as the health care system, schools, and government agencies whose decisions have significant downstream impacts. Families are viewed by many public-level entities as a reference when allocating resources, enforcing new standards, and responding to emerging needs and demands. From this perspective, the family operates at the nexus of individual and societal structures, embedding the tripple effects of public policy in its most consequential ways. Finally, the family has a crucial role to play in determining the life course of its members. As COVID-19 elevates the importance of familial bonds, the post-pandemic period might be marked by an increased influence of families on the lives of their members and the social structures they inhabit.

The above conceptualization echoes several other approaches to studying long-term health inequalities. For instance, life course theory emphasizes the significance of family ties and social environment as they structure prospects for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan, 2005; Settersten et al., 2020). Likewise, social capital theory and the stress-buffering hypothesis consider family systems a crucial factor in health and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills, 1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987; Mandelbaum et al., 2020). They show that the family can offer tangible (e.g., food, shelter, and money) and intangible resources (e.g., support, social status, psychological comfort, and institutional advantages), protecting its members from adverse life events. Similarly, the social-ecological model and the theory of fundamental causes highlight how the SES, familial and extrafamilial factors pattern people's life trajectories and health outcomes (Barkan and Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et al., 2012; Ungar, 2015). However, all these approaches are ostensibly focused on life periods where the general social context is relatively stable. In this article, we present our conceptualization of the family in times of unprecedented social upheaval. Specifically, our framework seeks to characterize family dynamics as a suitable lens to analyze health inequalities in the face of a virulent public health crisis.

# Methodology

## Data

The data for this study come from an original collection of biographical data on immigrants and the native population of Quebec to analyse their participation in Quebec society in various fields (social, economic, cultural, etc.). The immigrant population is limited to recent immigration that has taken place over the past 15 years. This survey has several components that retrace the life course of individuals over a long period of time. The pandemic component uses a reduced version of the [harmonized questionnaire on exposure and the impact of Covid19 on the family (CEFIS)](https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf). The questions are divided into three parts, on exposure to the disease, its consequences and the benefit of the policies put in place by both the federal and provincial governments. The other parts of the questionnaire focus on the characteristics of the respondents, their immigration history, their educational and family background. In short, it is a detailed questionnaire that covers the life of immigrants from birth and especially over the past 15 years.
The data collection lasted three months, from August to November 2020. In total, 1,500 people were surveyed, including around 1,000 immigrants and 500 natives throughout Quebec.

### Dependent variables

We distinguish three groups of dependent variables:

**1.	Disease exposure variables**: these are measured by six questions that seek to document whether the individual through work or a family member may have been exposed to the disease. For example, it is asked if a family member has had symptoms of the disease, or if a member has been hospitalized due to the contraction of the disease.

**2.	Variables on the family consequences of the disease**: we measure three dimensions of consequences. The first group of measures assess the consequences on accessing ressources by the respondent or any member of his family. The second group of measure assess whether the respondent or any family member have difficulties to perform usual duties toward any member of the family. Finally, the last group of measures measure the consequences of the pandemic on employment and income of any family member.

**3.	Variables on government assistance**: Finally, government support is measured by the following two questions that assessed whether the respondent or any family member has benefited of the temporary assistance for workers from the Government of Quebec (PATT) or of the Canadian Federal Government Emergency Benefit (CEP). The table XXX in the apendix presents the full set of indicators of the dependents variables

### Main independent variable
The main independent variable is immigrant status coupled with visible minority status. This variable takes the following modalities: 1 if the respondent is a non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a racialized immigrant, and 4 if the respondent is a racialized native or an immigrant who has not answered the visible minority question. The other variables in the analysis are: age, sex, length of residence, minority status, size of social network.

## Analysis methods

We use two main methods of analysis, a descriptive method and an explanatory method. In the descriptive method, we consider each of the independent variables in isolation and describe how they vary depending on the main independent variable. The explanatory methods goes further by assessing whether the effect remain when controlling for other covariates. Thus, by controlling for the other explanatory variables, we determine the effect of immigrant status on exposure to the pandemic, on its consequences and on the fact of benefiting or not of government assistance. In a second series of modelling, we assess whether the effect that we observed varies by the immigrant's length of residence or by marital status. We use either a logistic regression on the outcome variables that are binary and an ordonned logistic regression for the outcome that are measure with a 5-point Lickert scale. 
<!-- PUT THIS AS A LIMITATION IN THE CONCLUSION
To be able to benefit from government benefits, one must meet a number of conditions, including the loss of job due to Covid-19. As a result, the beneficiaries are not the general population but those who have lost their jobs and this may itself be due to immigrant status. Thus, for model 3, we will use Heckman's models by first estimating the fact of having lost or not a job and then, the fact of receiving or not the government benefit.
-->

# 3. Results

## 3.1. Sample distribution

## 3.2. Exposure by immigration status

Figure 1 presents the exposure to the pandemy accroding to the immigrants and minority visible status. For all the six measures of exposure, it appears that the majority group is the most exposed to the COVID-19 in all instance. The big difference between the groups is about the exposure trhough working outside. Around 50% of the member of this group or a family member work outside the home at the time of the survey. This percentage is only 30% for the racialised immigrants and 25% for the non-racialised immigrants.

![Figure 1.1: Exposure](Figures/Exposition1.png){width=60%}

![Figure 1.2: Explicative model](Figures/Exposure.jpeg)

## 3.3. Consequences by immigrations status

The Figure 2 presents the consequences of the COVID-19 for a family member to access to ressources. It shows that in all accounts but one (getting medecine), the racialised immigrants express more diffciult to assess ressources. Around 20% of them stated having difficulty to get health care compare, around 10% for obtaining food and 7% for obtaining family insurrance. Comparatively, these proportion are respectively 13%, 7% and 4% for people and thier family member in the majority group and 19%, 3% and 4% respectively among the racialised immigrants.

![Figure 3.1: Consequences on accessing ressources](Figures/Famille1.png){width=80%}

![Figure 3.2: Explicative model](Figures/Ressources.jpeg)

## 3.4. Consequences on performing usual duties

The Figure 4 compares the response from the three groups on performing usual duties. It appears that for all groups the situation caused by the COVID-19 has inexpectedly improved the situation in various way, irrespective of the immigrants status of the respondents.

![Figure 4.1: Consequences on performing usual duties](Figures/Entente1.png)

![Figure 4.2: Explicative model](Figures/Abilities.jpeg)

## 3.5. Consequences on employment by immigration status

The consequences on employment are mixed, with the severe consequences more prevalent among the racialised immigrants. The Figure 3 shows for instance that close to 45% of the racialised immigrants have stated  that their family income has decreased compare to just 33% of the majority group and 37% of the non-racialed immigrants. In addition, 11% of the racialised immigrants reveals having loss their job compared to 9% among the non-racialised immigrants and 2% among members of the majority group. At the opposite, the members of the majority groups are more likely to report a reduction of the working hours or for stopping momentarily their work.

![Figure 5.1: Consequences on employment and income](Figures/Emploi1.png)

![Figure 5.2: Explicative model](Figures/Employment.jpeg)

## 3.7 Help received by immigrations status

Finally, we compared the different group on benefitting of the government assistance. It appears that racialised immigrants with 35% of the case have expressed having received the Canada Emergency Response Benefit wheras this proportion is 31% for the non-racialised immigrants and 25% for the member of the majority group. For the province Temporary Aid for Workers Program, around 5% of the respondents have received it.

![](Figures/Aide1.png)

![Figure 3.2: Explicative model](Figures/Government.jpeg)

# Conclusion

The preliminary results of this study show that the consequences of the pandemic on the Quebec population differ according to the population groups, but also according to the type of consequences.

>>>>

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---

@@
